Discrete Quantum Theories and Computing

### 2018-11-26 dissertation.tex
dissertation.lyx "2.2.2 Gleason's Theorem and The Born Rule" Fix typos...

%% LyX 2.3.1-1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,twoside]{iuphd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{Times New Roman}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{verbatim}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage[all]{xy}

@@

\usepackage{polyglossia}
\setdefaultlanguage[variant=american]{english}
\setotherlanguage{arabic}
\usepackage[style=numeric-comp,sorting=none,maxbibnames=10,doi=false,isbn=false,url=false]{biblatex}
\addbibresource{prop.bib}
\begin{document}
\title{Discrete Quantum Theories and Computing}
\author{Tai, Yu-Tsung}

@@

\begin{acknowledgments*}\todo{Confirm people to be acknowledged
like the way being acknowledged... }I want to thank Prof.~Gerardo
Ortiz, Prof.~Amr Sabry (\textarabic{عمرو صبري}), and Prof.~Andrew
Hanson for publishing three papers \cite{geometry2013,DQT2014,THOS2018}
with me, and enormous other things. These papers become the main part
of my thesis, especially most of Chapter~\ref{chap:DQT=000026DQC}
is based on “Geometry of discrete quantum computing”~\cite{geometry2013}
and “Discrete quantum theories” (DQT)~\cite{DQT2014}; most
of Chapter~\ref{chap:QIVPM} is based on “Quantum Interval-Valued
Probability: Contextuality and the Born Rule” (QIVPM) \cite{THOS2018}.
Together with Prof.~Dylan Thurston and Prof.~Shouhong Wang (汪守宏),
I want to thank you for getting together to understand what I did
and passing my dissertation proposal Spring 2017. I also thank Prof.
Dylan Thurston for serving in my advisory committee and my Tier 3
committee, with other advisory committee members passing my Computer
Science qualifying exam Spring 2016. Especially, my survey of real
computation in the qualifying exam inspired me the comparison of quantum
computer and analog computer in Chapter~\ref{chap:Introduction}.
Chapter~\ref{chap:Introduction} also contains the motivation in
my application for Rethinking Foundations of Physics 2017 workshop
reviewed by Prof. Amr Sabry (\textarabic{عمرو صبري}). I want
to thank Prof. Lawrence Moss for inviting me to present our results
in the interdisciplinary logic seminar and theory seminar in Computer
Science, co-chairing my Tier 3 committee, writing an assessment letter
for me, and drawing my attention on Prof.~Abramsky's paper~\cite{Abramsky2012},
which inspired me to merge quantum probability with discrete quantum
theories. And I thank John Gardiner for inspiring discussion on the
difficulties merging them~\cite{Gardiner2014}. These results are
improved and motivated the transition from DQT to QIVPM in Sec.~\ref{sec:Toward-IVPM}.
I also thank Prof.~Tom Lewis for helping me organize the unpublished
deterministic quantum algorithms in the term paper for SLST-T501 Academic
Writing course as presented in Sec.~\ref{subsec:Discrete-Deutsch-algorithm}.
I want to thank Traci Nagle for going through the literature review
about finite precision measurement and contextuality discussed for
CSCI-Y790 Writing and Editing course as presented in Sec.~\ref{sec:Quantum-Contextuality}.
I also thank Elizabeth “Betsy” Merceron and Kexin “Casey”
Chen (谌可心) for teaching me English in general so that others can understand
me more easily.

@@

Let a pure unnormalized state $\ket{\Phi}\in\mathcal{H}$ characterize
the particle sending to a Stern-Gerlach type experiment, and $\muB_{\Phi}$
be the quantum probability measure of the resulting events sensibly
corresponding to $\ket{\Phi}$. For a correspondence $\ket{\Phi}\mapsto\muB_{\Phi}$
to be sensible, we hope that if the state~$\ket{\Phi}$ is one of
the outcomes of a quantum event~$P$, $P\ket{\Phi}=\ket{\Phi}$,
then the event~$P$ always happens, $\muB_{\Phi}\left(P\right)=1$,
and vice versa. Moreover, since the physical phenomena exist and should
be the same no matter how we describe them, the probability of an
event should be invariant despite how we choose the basis. Because
changing to another basis is the same as applying an unitary map~$U$,
we should have $\muF_{U\ket{\Phi}}\left(UPU^{\dagger}\right)=\muF_{\Phi}\left(P\right)$.
It is easy to check that the correspondence satisfying the conditions
in this paragraph is unique, 
\begin{equation}
\muB_{\Phi}(P)=\frac{\melem{\Phi}{P}{\Phi}}{\ip{\Phi}{\Phi}}\label{eq:Born}
\end{equation}
and called the Born rule \cite{Born1983,Mermin2007,Jaeger2007}. If
$\ket{\Phi}$ is normalized, Eq.~(\ref{eq:Born}) could be simplified
as $\muB_{\Phi}(P)=\melem{\Phi}{P}{\Phi}$. Since a mixed state $\rho=\sum_{j=1}^{N}q_{j}\proj{\Phi_{j}}$
is a weighted average of projectors $\proj{\Phi_{j}}$ with weights
$q_{j}>0$ and $\sum_{j=1}^{N}q_{j}=1$, the generalized Born rule
of $\rho$, $\muB_{\rho}\left(P\right)$, is also a weighted average
of $\muB_{\Phi_{j}}\left(P\right)$, 
\begin{equation}
\muB_{\rho}\left(P\right)=\sum_{j=1}^{N}q_{j}\muB_{\Phi_{j}}\left(P\right)=\Tr\left(\rho P\right)\,,\label{eq:mixed-Born}
\end{equation}
which is also a quantum probability measure and consistent with Gleason's
theorem.

>>>>

%% LyX 2.3.1-1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[11pt,twoside,english]{iuphd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{Times New Roman}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{verbatim}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{enumitem}
\usepackage{pdfpages}
\usepackage{graphicx}
\usepackage[all]{xy}

@@

\usepackage{polyglossia}
\setdefaultlanguage[variant=american]{english}
\usepackage[style=numeric-comp,sorting=none,maxbibnames=10,doi=false,isbn=false,url=false]{biblatex}
\addbibresource{prop.bib}
\begin{document}
\title{Discrete Quantum Theories and Computing}
\author{Tai, Yu-Tsung}

@@

\begin{acknowledgments*}\todo{Confirm people to be acknowledged
like the way being acknowledged... }I want to thank Prof.~Gerardo
Ortiz, Prof.~Amr Sabry, and Prof.~Andrew Hanson for publishing
three papers \cite{geometry2013,DQT2014,THOS2018} with me, and enormous
other things. These papers become the main part of my thesis, especially
most of Chapter~\ref{chap:DQT=000026DQC} is based on “Geometry
of discrete quantum computing”~\cite{geometry2013} and “Discrete
quantum theories” (DQT)~\cite{DQT2014}; most of Chapter~\ref{chap:QIVPM}
is based on “Quantum Interval-Valued Probability: Contextuality
and the Born Rule” (QIVPM) \cite{THOS2018}. Together with Prof.~Dylan
Thurston and Prof.~Shouhong Wang (汪守宏), I want to thank you for getting
together to understand what I did and passing my dissertation proposal
Spring 2017. I also thank Prof. Dylan Thurston for serving in my advisory
committee and my Tier 3 committee, with other advisory committee members
passing my Computer Science qualifying exam Spring 2016. Especially,
my survey of real computation in the qualifying exam inspired me the
comparison of quantum computer and analog computer in Chapter~\ref{chap:Introduction}.
Chapter~\ref{chap:Introduction} also contains the motivation in
my application for Rethinking Foundations of Physics 2017 workshop
reviewed by Prof. Amr Sabry. I want to thank Prof. Lawrence Moss
for inviting me to present our results in the interdisciplinary logic
seminar and theory seminar in Computer Science, co-chairing my Tier
3 committee, writing an assessment letter for me, and drawing my attention
on Prof.~Abramsky's paper~\cite{Abramsky2012}, which inspired me
to merge quantum probability with discrete quantum theories. And I
thank John Gardiner for inspiring discussion on the difficulties merging
them~\cite{Gardiner2014}. These results are improved and motivated
the transition from DQT to QIVPM in Sec.~\ref{sec:Toward-IVPM}.
I also thank Prof.~Tom Lewis for helping me organize the unpublished
deterministic quantum algorithms in the term paper for SLST-T501 Academic
Writing course as presented in Sec.~\ref{subsec:Discrete-Deutsch-algorithm}.
I want to thank Traci Nagle for going through the literature review
about finite precision measurement and contextuality discussed for
CSCI-Y790 Writing and Editing course as presented in Sec.~\ref{sec:Quantum-Contextuality}.
I also thank Elizabeth “Betsy” Merceron and Kexin “Casey”
Chen (谌可心) for teaching me English in general so that others can understand
me more easily.

@@

Let a pure unnormalized state $\ket{\Phi}\in\mathcal{H}$ characterize
the particle sending to a Stern-Gerlach type experiment, and $\muB_{\Phi}$
be the quantum probability measure of the resulting events sensibly
corresponding to $\ket{\Phi}$. For a correspondence $\ket{\Phi}\mapsto\muB_{\Phi}$
to be sensible, we hope that if the state~$\ket{\Phi}$ is one of
the outcomes of a quantum event~$P$, $P\ket{\Phi}=\ket{\Phi}$,
then the event~$P$ always happens, $\muB_{\Phi}\left(P\right)=1$,
and vice versa. Moreover, since the physical phenomena exist and should
be the same no matter how we describe them, the probability of an
event should be invariant despite how we choose the basis. Because
changing to another basis is the same as applying an unitary map~$U$,
we should have $\muB_{U\ket{\Phi}}\left(UPU^{\dagger}\right)=\muB_{\Phi}\left(P\right)$.
It is easy to check that the correspondence satisfying these conditions
is unique, 
\begin{equation}
\muB_{\Phi}(P)=\frac{\melem{\Phi}{P}{\Phi}}{\ip{\Phi}{\Phi}}\label{eq:Born}
\end{equation}
and called the Born rule \cite{Born1983,Mermin2007,Jaeger2007}. If
$\ket{\Phi}$ is normalized, Eq.~(\ref{eq:Born}) could be simplified
as $\muB_{\Phi}(P)=\melem{\Phi}{P}{\Phi}$. Since a mixed state $\rho=\sum_{j=1}^{N}q_{j}\proj{\Phi_{j}}$
is a weighted average of projectors $\proj{\Phi_{j}}$ with weights
$q_{j}>0$ and $\sum_{j=1}^{N}q_{j}=1$, the generalized Born rule
of $\rho$, $\muB_{\rho}\left(P\right)$, is also a weighted average
of $\muB_{\Phi_{j}}\left(P\right)$, 
\begin{equation}
\muB_{\rho}\left(P\right)=\sum_{j=1}^{N}q_{j}\muB_{\Phi_{j}}\left(P\right)=\Tr\left(\rho P\right)\,,\label{eq:mixed-Born}
\end{equation}
which is also a quantum probability measure and consistent with Gleason's
theorem.
